---
title: "TidyTuesday - Bechdel Test"
output: html_notebook
---

## TidyTuesday - 9 March 2021
The dataset is loaded from the tidytuesdayR package

```{r import_data_packages}
library(tidytuesdayR)
library(summarytools)
library(tidyverse)
library(dplyr)

# Load data
tuesdata <- tidytuesdayR::tt_load('2021-03-09')
tuesdata <- tidytuesdayR::tt_load(2021, week = 11)

bechdel <- tuesdata$movies

head(bechdel)

```

### The dataset
The dataset contains 1,794 observations and 23 variables

```{r viewdataset}
library(summarytools)

view(dfSummary(bechdel), method = 'render')

```
### Selecting a dataset for analysis

The following variables will be retained for analysis:

- year

- title

- clean_test

- binary

- budget_2013

- domgross_2013

- intgross_2013

- period_code

- decade_code

- rated

- metascore

- imdb_rating

- genre

```{r select_short}
bechdel_short <- bechdel %>%
  select(year, title, clean_test, binary, budget_2013, domgross_2013, intgross_2013, period_code, decade_code, rated, metascore, imdb_rating, genre)
```

### Investigate `clean_test` variable

- The values: "dubious", "men", "notalk" and "nowomen" all represent Bechdel test failures

- "ok" represents a Bechdel test pass

- The variable needs to be converted into a factor with the following order:
  - 1. No women
  - 2. No talk
  - 3. Men
  - 4. Dubious
  - 5. OK

```{r clean_test}

clean_test_levels <- c("nowomen", "notalk",
                       "men", "dubious", "ok")

bechdel_short <- bechdel_short %>%
  mutate(clean_test = factor(clean_test, levels = clean_test_levels))

ggplot(bechdel_short,
       aes(x = clean_test, fill = binary)) +
  geom_bar()

```

### Relationship between clean_test and year
The number of observations significant increases after 1995, so its difficult to tell how the distribution changes by year

```{r}
ggplot(bechdel_short, aes(x = year, fill = clean_test)) +
  geom_histogram(binwidth = 5)
```

### Understanding period codes

- Period 1 = 2010-2013 (c. 400 obs)

- Period 2 = 2005-2009 (c. 500 obs)

- Period 3 = 2000-2004 (c. 350 obs)

- Period 4 = 1995-1999 (c. 250 obs)
 
- Period 5 = 1990-1994 (c. 90 obs)

- Period NA = 1970-1989 (c. 180 obs)

```{r}
bechdel_short %>%
  select(period_code, year) %>%
  group_by(period_code) %>%
  summarize(min_year = min(year), max_year = max(year), count = n())

```
### Relationship between clean_test and period_code

- The proportion of films that passed the BT increased from period 5 (1990 - 1994) to period 3 (2000 to 2004), after which the proportion has been declining

```{r period_code}
ggplot(bechdel_short, aes(x = period_code, fill = clean_test)) +
  geom_bar(position = "fill")
```

### Relationship between clean_test and budget_2013
- Films that pass the Bechdel Test tend to have slightly lower budgets than those that don't - particularly those that fewer than 2 women, or where women don't talk to each other

```{r budget_2013}
ggplot(bechdel_short, aes(x = clean_test, y = budget_2013)) +
  geom_boxplot() +
  scale_y_log10()
```

### Relationship between clean_test and domgross_2013
- There doesn't appear to be obvious differences between BT categories and domestic gross in 2013 currency

```{r domgross_2013}
ggplot(bechdel_short, aes(x = clean_test, y = as.numeric(domgross_2013))) +
  geom_boxplot() +
  scale_y_log10()
```

### Relationship between clean_test and intgross_2013
- There doesn't appear to be obvious differences between BT categories and international gross in 2013 currency

```{r intgross_2013}
ggplot(bechdel_short, aes(x = clean_test, y = as.numeric(intgross_2013))) +
  geom_boxplot() +
  scale_y_log10()
```

### Relationship between clean_test and rated (where rated is PG, PG-13, or R)
- Films that are PG or PG-13 are more likely to pass or closely pass (dubious) the Bechdel Test

```{r rated}
bechdel_short_rated <- bechdel_short %>%
  filter(rated %in% c("PG", "PG-13", "R"))

ggplot(bechdel_short_rated, aes(x = rated, fill = clean_test)) +
  geom_bar(position = "fill")

```

### Relationship between clean_test and metascore
- There is no clear relationship between `clean_test` and `metascore`

```{r metascore}
ggplot(bechdel_short, aes(x = clean_test, y = metascore)) +
  geom_boxplot()
```

#### Relationship between clean_test and imdb_rating
- There some evidence that films that pass or almost pass ("dubious") the Bechdel Test have a lower imdb rating

```{r imdb_rating}
ggplot(bechdel_short, aes(x = clean_test, y = imdb_rating)) +
  geom_boxplot()
```

### Use of the genre variable
- This variable needs further cleaning before it can be used in analysis

- The column contains multiple genres separated by commas

- Max number of genres = 3

```{r genre}
bechdel_short_genres <- bechdel_short %>%
  group_by(genre) %>%
  summarize(count = n())

bechdel_short_genres$commas <- str_count(bechdel_short_genres$genre, pattern = ",")

bechdel_short_genres %>%
  group_by(commas) %>%
  summarize(max(commas))

bechdel_short_genres
```
### The relationship between year, imdb_rating and Bechdel test

```{r year_rating}
ggplot(bechdel_short, aes(x = year, y = imdb_rating, color = binary)) +
  geom_point(position = "jitter") +
  geom_smooth(method = "lm")
```

### Split genres into columns
- To compare relationships between genres, each genre observation (which can contain multiple genres) has been split into a separate observation

- This means that some films are contained in the table multiple times, however the maximum number of genres for any given film is 3. As a result, I don't expect this to result in significant bias

```{r genre_split}

# Split genre into 3 columns
bechdel_genre_split <- separate(bechdel_short, 'genre', 
                                paste('genre', 1:3, sep="_"), 
                                sep=",", extra = "warn")

# Create a separate dataframe for genre_1
genre_1 <- bechdel_genre_split %>%
  select(-genre_2, -genre_3) %>%
  mutate(genre = str_trim(genre_1, side = "both")) %>%
  select(-genre_1)

genre_1

# Create a separate dataframe for genre_2
genre_2 <- bechdel_genre_split %>%
  select(-genre_1, -genre_3) %>%
  mutate(genre = str_trim(genre_2, side = "both")) %>%
  select(-genre_2)

# Create a separate dataframe for genre_3
genre_3 <- bechdel_genre_split %>%
  select(-genre_1, -genre_2) %>%
  mutate(genre = str_trim(genre_3, side = "both")) %>%
  select(-genre_3)

# All tables combined
all_genres <- rbind.data.frame(genre_1, genre_2, genre_3)

all_genres

# Calculate the max, min and median rating by genre, by pass/fail
genre_data_all <- all_genres %>%
  group_by(genre) %>%
  summarize(max_rating = max(imdb_rating), 
            min_rating = min(imdb_rating), 
            med_rating = median(imdb_rating),
            avg_rating = mean(imdb_rating)) %>%
  arrange(desc(avg_rating)) %>%
  mutate(genre = factor(genre, levels = genre))

genre_data_pass <- all_genres %>%
  group_by(genre, binary) %>%
  summarize(max_rating = max(imdb_rating), 
            min_rating = min(imdb_rating), 
            med_rating = median(imdb_rating),
            avg_rating = mean(imdb_rating)) %>%
  filter(binary == "PASS") %>%
  arrange(desc(avg_rating))

genre_data_fail <- all_genres %>%
  group_by(genre, binary) %>%
  summarize(max_rating = max(imdb_rating), 
            min_rating = min(imdb_rating), 
            med_rating = median(imdb_rating),
            avg_rating = mean(imdb_rating)) %>%
  filter(binary == "FAIL") %>%
  arrange(desc(avg_rating))

```

```{r genre_chart_all}
base_plot <- ggplot(genre_data_all, aes(x = genre)) +
  geom_segment(aes(x = genre, 
                   xend = genre, 
                   y = min_rating, 
                   yend = max_rating),
               color = "azure3") +
  geom_point(aes(x = genre, y = min_rating), 
             color = "firebrick3", size = 3) +
  geom_point(aes(x = genre, y = max_rating), 
             color = "aquamarine4", size = 3) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.25, hjust = 1)) +
  ylab("Minimum and Maximum IMDB rating") +
  xlab("Genre")

base_plot

  
```

```{r chart_data}
# Calculate the max, min and median rating by genre, by pass/fail
genre_data_chart <- all_genres %>%
  group_by(genre) %>%
  filter(genre != "NA") %>%
  summarize(max_rating = max(imdb_rating),
            min_rating = min(imdb_rating),
            avg_pass_rating = mean(imdb_rating[binary == "PASS"]),
            avg_fail_rating = mean(imdb_rating[binary == "FAIL"]),
            med_pass_rating = median(imdb_rating[binary == "PASS"]),
            med_fail_rating = median(imdb_rating[binary == "FAIL"]),
            avg_rating = mean(imdb_rating),
            med_rating = median(imdb_rating)) %>%
  arrange(desc(avg_rating)) %>%
  mutate(genre = factor(genre, levels = genre))
  

genre_data_chart
```

```{r generate_chart}
genre_chart <- ggplot(genre_data_chart, aes(x = genre)) +
  # Create the line plots
  geom_segment(aes(x = genre, 
                   xend = genre, 
                   y = min_rating, 
                   yend = max_rating),
               color = "gray85") +
  # Add a point for min rating
  geom_point(aes(x = genre, y = min_rating), 
             color = "paleturquoise3", size = 3) +
  # Add a point for max rating
  geom_point(aes(x = genre, y = max_rating), 
             color = "paleturquoise3", size = 3) +
  # Add a line for the median rating for BT pass films
  geom_line(aes(y = med_pass_rating, group = 1),
            color = "aquamarine4", size = 1) +
  # Add a line for the median rating for BT fail films
  geom_line(aes(y = med_fail_rating, group = 1),
            color = "firebrick3", size = 1) +
  
  # Set light theme
  theme_light() +
  # Rotate x axis labels 90 degrees
  theme(axis.text.x = element_text(angle = 90, vjust = 0.25, hjust = 1)) +
  # Remove gridlines
  theme(panel.grid.major = element_blank()) +
  theme(panel.grid.minor = element_blank()) +
  # Add labels for x and y axis
  labs(x = "Genre", y = "IMDB rating", 
       title = "Film ratings by Genre \n...and whether they pass the Bechdel Test") +
  # Add labels for min and max ratings
  annotate("text", x = "Music", y = 10, 
           label = "Max rating", size = 3) +
  annotate("text", x = "Music", y = 2,
           label = "Min rating", size = 3) +
  # Add labels for median ratings for BT pass and fail films
  annotate("text", x = "Comedy", y = 7.25,
           label = "Median rating - Bechdel fail",
           size = 3) +
    annotate("text", x = "Comedy", y = 5.9,
           label = "Median rating - Bechdel pass",
           size = 3)

genre_chart

ggsave("tidytues_bechdel.png")

```

